---
title: "Benchmarking DE Tests"
author: 
  - "Rance Nault"
  - "Satabdi Saha"
package: "statsDR"
data: "Last updated: April 12, 2021"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
  vignette: >
    %\VignetteIndexEntry{simple example}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r}
# Load packages
## TODO: Reduce requirements - look for redundancy
suppressPackageStartupMessages({
  library(snseq.stats)
})
```
# DETesting

- List all the tests that are available (ANOVA, KW, WRS, _LRT_linear_, _LRT_multiple_, _ChiSQ_, MAST, Bayes). Describe the italicized ones. Mention that we updated the test (MAST, WRS) to combined all pairwise. These tests can be run using `DETest`.

## LRT_linear
Describe
```{r}
LRT.linear = DETest(sce, method = 'LRT.linear')
head(LRT.linear)
```

## LRT_multiple
```{r}
LRT.multiple = DETest(sce, method = 'LRT.multiple')
head(LRT.multiple)
```

## ChiSQ
```{r}
chi.square = DETest(sce, method = 'CHISQ')
head(chi.square)
```

How to list the options of method #TODO: Check how people do this, if they do. 
```{r}

```


#Last
```{r}
All = DETest(sim, method = 'All') #This will take a long time depending on the amount of genes.
#The output should be a list of dataframes that have a p-value/sig value
names(All)
```


```{r}
significantGenes = getDEGs(sim, All, threshold = 0.05)
```

```{r}
merged.df = Reduce(merge, lapply(list(KW, fc.max, pz.min), function(x) data.frame(x, rn = row.names(x))))
colnames(merged.df) = c("Gene", "pvalue", "fc.max", "pz.min")
filtered = merged.df %>% filter(pvalue <= threshold & fc.max >= fc.threshold & pz.min >= pct.expressed)
```


```{r}
new_bayesDETest(h)
```


# Benchmarking
1) Extract p-values of all the test
Data.frame, allgenes = row, column = test for all genes that meet a criteria. Make one for each test 

How to handle pairwise tests -- pair matrices/foldchange by index
```{r}
getDEGs = function(DETestoutput, threshold = 0.05, bayes.threshold = 1/3, fc.threshold = 1, pct.expressed = 0){
  
  
  for (test in DETestoutput){
    if (names(test) != bayes){
      #apply 0.05 filter
    } else {
      #expbf < 1/3
    }
  }
  return a dataframe
}
```


2) EXtract 'truth'
```{r}

```











