---
title: "Differential gene expression analysis for dose dependent simulated data"
output: html_notebook
---

Data Description: 

simulated.data is simulated lognormcounts matrix having 2300 genes and 4500 cells. 
simulated.cell.metadata is the cell metadata matrix having the dose level information for the 4500 cells. There are a total of 9 dose groups with 500 cells each.
simulated.gene.metadata is the gene metadata matrix having 2300 rows and 4 columns, where the columns correspond to gene names,model: the model dose response curve that the gene follows and Direction: which states whether the gene is UpRegulated or DownRegulated.

```{r}
# simulated.data <- read.delim("~/Documents/Documents/Summer_2019/Mouse_Liver/Rance/snSeq-DGE/New_Simulations/simDataFinal.lognormcounts.txt")
# rownames(simulated.data)<-simulated.data[,1]
# simulated.data<-simulated.data[,-1]
# simulated.cell.metadata <-read.delim("~/Documents/Documents/Summer_2019/Mouse_Liver/Rance/snSeq-DGE/New_Simulations/simCellMeta.txt")
# simulated.gene.metadata <- read.delim("~/Documents/Documents/Summer_2019/Mouse_Liver/Rance/snSeq-DGE/New_Simulations/simGeneMeta.txt")
# simulated.gene.metadata<-simulated.gene.metadata[,-1]
```

Data Description: 

simulated.data is simulated lognormcounts matrix having 2500 genes and 4500 cells. 
simulated.cell.metadata is the cell metadata matrix having the dose level information for the 4500 cells. There are a total of 9 dose groups with 500 cells each.
simulated.gene.metadata is the gene metadata matrix having 2500 rows and 4 columns, where the columns correspond to gene names,model: the model dose response curve that the gene follows and Direction: which states whether the gene is UpRegulated or DownRegulated.

```{r}
simulated.data<-read.csv("~/Documents/Documents/Summer_2019/Mouse_Liver/Rance/snSeq-DGE/simulatedCorrectProportions/norm_sim_correctprop.txt", sep="")
simulated.cell.metadata<-read.csv("~/Documents/Documents/Summer_2019/Mouse_Liver/Rance/snSeq-DGE/simulatedCorrectProportions/sim_correctprop_cellMeta.txt", sep="")
simulated.gene.metadata<-read.csv("~/Documents/Documents/Summer_2019/Mouse_Liver/Rance/snSeq-DGE/simulatedCorrectProportions/sim_correctprop_geneMeta.txt", sep="")
```

```{r}
library(data.table)
library(ggplot2)
simulated.data_df <-as.data.table(t(simulated.data))
k0<-qplot(gene1, data = simulated.data_df[gene1  < 
                    quantile(as.numeric(simulated.data_df$gene1), probs = .95)]) 
simulated.data_df<- simulated.data_df[, gene1_pos := ifelse(gene1 > 0, gene1 , NA)]
k1<-ggplot(simulated.data_df, aes(x = gene1_pos)) + 
  geom_histogram(aes(y = ..density..), colour = "black", fill = "White") +
  stat_function(fun = dnorm, args = MASS::fitdistr(simulated.data_df[gene1 > 0 ,gene1_pos ], "normal")$estimate)
ggpubr::ggarrange(k0,k1,ncol=2,nrow=1) 

```

```{r}
dose_level<-factor(simulated.cell.metadata$doses)
simulated.data_dose<-data.frame(t(simulated.data),
                                              simulated.cell.metadata[,"doses"])
colnames(simulated.data_dose)[ncol(simulated.data_dose)]<-"Dose"
```

```{r}
data<-rep(list(data.frame()), nlevels(dose_level))
names(data)<-levels(dose_level)
for(i in levels(dose_level))
{
data[[i]]<- subset(simulated.data_dose
                   , Dose == i, select = gene1 : gene2300)
##data[[i]] is a list of all the seperate dose gene*cell matrices
data[[i]][] <- lapply(data[[i]], function(x) as.numeric(as.character(x)))
data[[i]]<-t(data[[i]])
}
```

```{r}
# Identify the genes expressed in less than 5% of the cells over all the dose groups
propor<-sapply(data,function(x) apply(x,1,function(x) length(which(x!=0))/length(x)))
b<-which(apply(propor,1,function(x) all(x>0.05))== TRUE)
#Filter out genes
data<-lapply(data,function(x) apply(x,2,function(x) x[b]))
simulated.gene.metadata<-as.data.frame(simulated.gene.metadata[b,])
```


Create data for applying LRT_multiple_groups
```{r}
data_short<-rep(list(data.frame()),length(data))
data_ind<-lapply(data, function(x) ifelse(x>0,1,0))
#data_ind_short<-lapply(data_ind, function(x) x[x!=0])
data_short<-lapply(data, function(x) apply(x,1,function(x) x[x!=0]))
gene_names<-vector()
for(j in 1:nrow(data[["0"]]))
{
  gene_names[j]<-sprintf("gene%i",j)
}
out<-rep(list(list()),times=nrow(data[["0"]]))
for(j in 1:nrow(data[["0"]]))
{
    xx<-noquote(sprintf("gene%i",j))
    l1<-list(data_short[["0"]][j][[xx]],
             data_short[["0.01"]][j][[xx]],data_short[["0.03"]][j][[xx]],
                      data_short[["0.1"]][j][[xx]],
                      data_short[["0.3"]][j][[xx]],data_short[["1"]][j][[xx]],
                          data_short[["3"]][j][[xx]],data_short[["10"]][j][[xx]],
                                  data_short[["30"]][j][[xx]])
    l2<-list(data_ind[["0"]][j,],data_ind[["0.01"]][j,],
                                  data_ind[["0.03"]][j,],data_ind[["0.1"]][j,],
                                  data_ind[["0.3"]][j,],data_ind[["1"]][j,],
                                  data_ind[["3"]][j,],data_ind[["10"]][j,],
                                  data_ind[["30"]][j,])
    out[[j]]<-LRT_multiple_groups(l1,l2)
    
}

```


Prior selection under alternative hypothesis

```{r}
library(matrixStats)
m<-vector()
sigma_mean<-vector()
sigma_var<-vector()
omega_mean<-vector()
omega_var<-vector()
for(i in  levels(dose_level))
{
  #Dose group specific mean of expression levels for all genes
  m[i]<-mean(rowMeans(replace(data[[i]], data[[i]] == 0, NA), na.rm = TRUE),na.rm = TRUE)
  #Dose group specific SD's for all genes
  sigma_mean[i]<-mean(rowVars(replace(as.matrix(data[[i]]),as.matrix(data[[i]]) == 0, NA), na.rm = TRUE),na.rm = TRUE)
  sigma_var[i]<-var(rowVars(replace(as.matrix(data[[i]]),as.matrix(data[[i]]) == 0, NA), na.rm = TRUE),na.rm = TRUE)
  #Dose group specific mean dropout proportions for all genes
  omega_mean[i]<-mean(apply(data[[i]],1,function(x) length(which(x==0))/length(x)))
  omega_var[i]<-var(apply(data[[i]],1,function(x) length(which(x==0))/length(x)))
}
#Choose the scaling constant tau_mu
tau_t_mu = tau_c_mu = tau_mu = c(0.5,1,2,3)
tau_k_mu<-rep(1,9)
#Use sigma_mean and sigma_var to solve for b_sigma and a_sigma
b_sigma = 1
a_sigma = 6
#Use mean and variance of omega to solve for a_w and b_w
#We keep the same values of a_w and b_w for all groups since the omega mean and omega SD values are more or less consistent.
a_w = a_t_w = 0.8 #t refers to the treatment groups
b_w = b_t_w = 0.2
#length(which(simulated.gene.metadata$Category == "Positive"))
prior_Alter<-c(1-((0.78)^(1/1800)),0.01,0.05,0.1, 0.2,0.3,0.4,0.5)
prior_Null<- 1-prior_Alter
```


```{r}
#actual bayes factor computed in overleaf
##BF FDR control using jeffreys 
a<-rep(list(list()),nrow(data[["0"]]))
bf_actual_01<-rep(list(a), nlevels(dose_level)-1)
names(bf_actual_01)<-levels(dose_level)[-1]
source("Bayes_factor.R")
for (i in levels(dose_level)[-1])
{
  for(j in 1:nrow(data[["0"]]))
  {
    bf_actual_01[[i]][[j]]<-Bayes_factor(Y_t=as.matrix(data[[i]][j,]),
                           Y_c=as.matrix(data[["0"]][j,]),m_t0,m_c0,m_0,
                           tau_t_mu,tau_c_mu, tau_mu=tau_mu[2], 
                           b_sigma,a_sigma,a_w,a_c_w,a_t_w,b_w,b_t_w,b_c_w,
                           prior_alt = prior_Alter[1],
                           prior_null = prior_Null[1])
  }
}
```


```{r}
#actual bayes factor multiple computed in overleaf
##BF FDR control using jeffreys 
bf_multiple_01<-rep(list(list()), nrow(data[["0"]]))
source('~/Documents/Documents/Summer_2019/Mouse_Liver/Rance/snSeq-DGE/R/BayesFactorMultiple.R')

  for(j in 1:nrow(data[["0"]]))
  {
    bf_multiple_01[[j]]<-Bayes_factor_multiple(
                           Y = list(Y_1=as.matrix(data[["0"]][j,]), 
                           Y_2=as.matrix(data[["0.01"]][j,]),
                           Y_3=as.matrix(data[["0.03"]][j,]),
                           Y_4=as.matrix(data[["0.1"]][j,]),
                           Y_5=as.matrix(data[["0.3"]][j,]),
                           Y_6=as.matrix(data[["1"]][j,]),
                           Y_7=as.matrix(data[["3"]][j,]),
                           Y_8=as.matrix(data[["10"]][j,]),
                           Y_9=as.matrix(data[["30"]][j,])),
                           m_0=mean(m),m=m,K=9,
                           tau_k_mu =tau_k_mu,tau_mu=tau_mu[2], 
                           b_sigma=1,a_sigma=6,a_w=8,b_w=2,
                           prior_alt =prior_Alter[8],
                           prior_null = prior_Null[8])
  }

```


```{r}
#Decision for all group differential expression
l_bf_multiple_01<-vector(mode = "numeric", length = nrow(data[["0"]]))
exp_bf_multiple_01<-vector(mode = "numeric", length = nrow(data[["0"]]))
Decision_bf_multiple_01<-vector(mode = "numeric", length = nrow(data[["0"]]))
Decision_bf_multiple_01_new<-vector(mode = "numeric", length = nrow(data[["0"]]))
for (j in 1: length(bf_multiple_01))
{
  l_bf_multiple_01[j]<-bf_multiple_01[[j]]$l_Bayes_factor_01
  exp_bf_multiple_01[j]<-exp(l_bf_multiple_01[j])
  Decision_bf_multiple_01[j]<- ifelse(exp_bf_multiple_01[j] > 1, "Fail to Reject H_0",ifelse( exp_bf_multiple_01[j] <= 1 & exp_bf_multiple_01[j] > (1/3), "Weak evidence against H_0",ifelse(exp_bf_multiple_01[j] <= (1/3) & exp_bf_multiple_01[j] > (1/10), "Moderate evidence against H_0", ifelse(exp_bf_multiple_01[j] <= (1/10) & exp_bf_multiple_01[j] > (1/30), "Substantial evidence against H_0",ifelse(exp_bf_multiple_01[j]<= (1/30) & exp_bf_multiple_01[j]> (1/100), " Strong evidence against H_0", ifelse(exp_bf_multiple_01[j] <= (1/100) & exp_bf_multiple_01[j] > (1/300), "Very strong evidence against H_0", "Decisive evidence against H_0"))))))
  Decision_bf_multiple_01_new[j]<-ifelse(Decision_bf_multiple_01[j] == "Fail to Reject H_0" | Decision_bf_multiple_01[j] == "Weak evidence against H_0", "Negative", "Positive")
}
Decision_bf_multiple_01_new<-
    factor(Decision_bf_multiple_01_new,levels = c("Positive","Negative"))
True_decision<-ifelse(simulated.gene.metadata$Direction=="Downregulated" | simulated.gene.metadata$Direction == "Upregulated","Positive","Negative")
True_decision<-factor(True_decision, levels = c("Positive","Negative"))
c<-data.frame(exp_bf_multiple_01,Decision_bf_multiple_01_new,
                     True_decision)
colnames(c)<-c("BF_Test_Statistic","Pred_Decision","True_Decision")
rownames(c)<-rownames(data[["0"]])
confusionMatrix_bf_multiple_01<-caret::confusionMatrix(c$Pred_Decision,
         c$True_Decision)
confusionMatrix_bf_multiple_01

#Look at the test-statistic values of those genes which are classified as false negative
FN<-c$BF_Test_Statistic[which(c$True_Decision=="Positive" & c$Pred_Decision == "Negative")]
ind<-rownames(c)[which(c$True_Decision=="Positive" & c$Pred_Decision == "Negative")]
c[ind,]
```


```{r}
df<-data.frame(categories<-c("FN","FP","TN","TP"),
               values = c(109,0,1757,460))
ggplot(df, aes(y=values, x= categories)) + geom_bar(stat = "identity",fill = "steelblue")+ geom_text(aes(label=values),size=3,vjust = -0.5)+
            xlab("Errors")+ ylab("Values")+
             ggtitle("tau_mu = 1, P(H_a)= 0.002")+ 
              theme(plot.title = element_text(color="red", size=10, face="bold.italic"))

```


```{r}
#Decisions for two group differential expression
l_bf_act_01<-rep(list(vector(mode = "numeric",
                             length = nrow(data[["0"]]))), nlevels(dose_level)-1)
bf_act_01<-rep(list(vector(mode = "numeric",
                             length = nrow(data[["0"]]))), nlevels(dose_level)-1)
Decision_bf_act_01<- rep(list(vector(mode = "numeric",
                             length = nrow(data[["0"]]))), nlevels(dose_level)-1)
Decision_bf_act_01_new<- rep(list(vector(mode = "numeric",
                             length = nrow(data[["0"]]))), nlevels(dose_level)-1)

names(l_bf_act_01)<-levels(dose_level)[-1]
names(bf_act_01)<-levels(dose_level)[-1]
names(Decision_bf_act_01)<-levels(dose_level)[-1]
names(Decision_bf_act_01_new)<-levels(dose_level)[-1]

for (i in levels(dose_level)[-1])
{
  for(k in 1: nrow(data[["0"]]))
  {
  l_bf_act_01[[i]][k]<-bf_actual_01[[i]][[k]]$l_Bayes_factor_01
  }
  bf_act_01[[i]]<-exp(l_bf_act_01[[i]])
  Decision_bf_act_01[[i]]<-ifelse(bf_act_01[[i]] > 1, "Fail to Reject H_0",ifelse( bf_act_01[[i]] <= 1 & bf_act_01 [[i]] > (1/3), "Weak evidence against H_0",ifelse(bf_act_01[[i]] <= (1/3) & bf_act_01[[i]] > (1/10), "Moderate evidence against H_0", ifelse(bf_act_01[[i]] <= (1/10) & bf_act_01[[i]] > (1/30), "Substantial evidence against H_0",ifelse(bf_act_01 [[i]]<= (1/30) & bf_act_01 [[i]] > (1/100), " Strong evidence against H_0", ifelse(bf_act_01 [[i]] <= (1/100) & bf_act_01 [[i]] > (1/300), "Very strong evidence against H_0", "Decisive evidence against H_0")))))) 
Decision_bf_act_01_new[[i]]<-ifelse(Decision_bf_act_01[[i]] == "Fail to Reject H_0" | Decision_bf_act_01[[i]] == "Weak evidence against H_0", "Negative", "Positive")
Decision_bf_act_01_new[[i]] <-
                          factor(Decision_bf_act_01_new[[i]],levels = c("Positive","Negative"))

}
True_decision<-simulated.gene.metadata$Category
c<-rep(list(data.frame()),nlevels(dose_level)-1)
names(c)<-levels(dose_level)[-1]
confusionMatrix_bf_act_01<-rep(list(list()),nlevels(dose_level)-1)
names(confusionMatrix_bf_act_01)<-levels(dose_level)[-1]
for(i in levels(dose_level)[-1])
{
  c[[i]]<-data.frame(bf_act_01[[i]],Decision_bf_act_01_new[[i]],
                     True_decision)
  confusionMatrix_bf_act_01[[i]]<-caret::confusionMatrix(c[[i]]$Decision_bf_act_01_new..i..,
                                    factor(c[[i]]$True_decision, 
                                           levels = c("Positive","Negative")))
  colnames(c[[i]])<-c("BF_Test_Statistic","Pred Decision","True Decision")
  rownames(c[[i]])<-rownames(data[["0"]])
}
#Save the list as a csv file
#lapply(c, function(x) write.table( data.frame(x), 'test.csv'  , append= T, sep=',' ))
#capture.output(c, file = "clust_1_short_filtered_results.txt")
#capture.output(confusionMatrix_bf_act_01, file = "clust_1_short_filtered_results_confusionMatrix.txt")
save(c, file="simulated_data_results.Rdata")
confusionMatrix_bf_act_01
```

```{r}
# library
#Confusion Matrix Plot
library(ggplot2)
# create a dataset
dose <- c(rep(0.01 , 4) , rep(0.03 , 4) , rep(0.1 , 4) , rep(0.3 , 4) , rep(1 , 4),rep(3 , 4) ,
            rep(10 , 4),rep(30 , 4)  )
type <- rep(c("TP" , "FP" , "FN","TN") , 8)
value<-c(31,0,538,1757,43,0,526,1757,50,0,519,1757,66,0,503,1757,96,0,473,1757,312,0,257,1757,
         159,0,410,1757,530,0,39,1757)
cf <- data.frame(dose,type,value)
```

```{r}

```

